---
title: "Sigleo & Bosley Yaquina Bay Study"
author: "Katelyn Bosley"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: html_document
subtitle: Data Analysis Log
geometry: margin=1in
---

```{r,echo = F, warning = F, comment = F, message=F}

#setting global options for the analysis log
library(knitr)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=70),tidy=TRUE, fig.width=6,
fig.height=5, warning=FALSE, message=FALSE,fig.align = "center")

# When tables are made no comments
options(xtable.comment = FALSE)


# Load in packages needed for the analysis
load.libraries<-function(){
  library(ggplot2)
  library(scales)
  library(lme4)
  library(dplyr)
  library(GGally)
  library(reshape2)
  library(gridExtra)
  library(grid)
  library(ggplot2)
  library(lattice)
  library(kableExtra)
}
load.libraries()

#########################################################################
#build a custom theme for the plots
cust_theme<-
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        legend.background = element_rect(fill="white"),
        panel.border = element_rect(colour = "black"))+
  #theme(legend.title = element_blank())+
  theme(strip.text.x = element_text(size = 10, colour = "black", face="bold"))
###########################################################################

#set the working directory

setwd("C:\\Users\\Katelyn Bosley\\Desktop\\Sigleo_MS\\YB_water_quality\\Analysis")

```
##Part 1: Water Column Characteristics
 \ 
 \
 \
Figure 1: Plot showing with daily flow rate and the nitrate concentration/flux across the study period.
```{r,echo = F, warning = F, comment = F, message=F}

flow<-read.csv("YaquinaFlow98-02_clipped.csv")
flow$date<-as.Date(flow$date,format="%m/%d/%Y")

#determine the sampling dates
dat<-read.csv("Y_iso_data_fixed2.csv")
dat$Date<-as.Date(dat$Date,format="%m/%d/%Y")
samp.d<-unique(dat$Date)



#create a plot with both N flux and flow

f.plot<-ggplot(flow, aes(x=date)) +
  geom_line(aes(y=flow, colour="River Flow"),size=1)+
  geom_line(aes(y=actual.flux/100, colour="Nitrate Load"), size=1, alpha=0.7)+
  scale_y_continuous(sec.axis = sec_axis(~.*100,name=expression(paste("Nitrate Load (kg d"^-1," )" ))))+
  scale_colour_manual(name = "Measurement", values=c("blue","red")) +
  scale_x_date(breaks = date_breaks("3 months"),
  labels = date_format("%y-%b"))+
  geom_vline(dat=samp.d, xintercept=samp.d, col="grey50",lty="longdash",size=0.1)+
  labs(x="Date (Yr-Mo)",y=expression(paste("Flow (m"^3," s"^-1," )" )))+
  cust_theme+
  theme(axis.text.x = element_text(angle = 45,hjust=1))+
  theme(legend.position = c(0.88, 0.90))
  #theme(legend.position = "topright")
f.plot

#png('Flow_plot.png')
#f.plot
#dev.off()

```
\
Here I created an additional figure to look at all the water column data. I omitted the samples that were taken at odd locations (7.1, 6b, 5b etc).  
\
```{r,echo = F, warning = F, comment = F, message=F}
#set the working directory

# Fix up the station codes
dat$Station_alt<-as.factor(dat$Station_alt)
#unique(dat$Station_alt)

#select the 
mat.dat<-dat[,c(3,8,9,10,11,12,13,14)]

#generate pdf with plots
#pdf("YB_pairs.pdf") 
ggpairs(mat.dat)
#dev.off()
```
\
\
\
Next, I divided the data into seasons for comparison of water column characteristics. This allows us to boost sample numbers and generate estimates of variance. \
\
**Winter** = Dec, Jan, Feb\
**Spring** = March, Apr, May\
**Summer** = June, July, Aug\
**Fall** = Sept, Oct, Nov
```{r,echo = F, warning = F, comment = F, message=F}
#divide the data by season
dat$Season=NA

for(i in 1:nrow(dat)){
if(dat$Month[i]=="12" || dat$Month[i]=='1'||dat$Month[i]=='2') {dat$Season[i]="Winter"}
  
if(dat$Month[i]=="3" || dat$Month[i]=='4'||dat$Month[i]=='5') {dat$Season[i]="Spring"}
  
if(dat$Month[i]=="6" || dat$Month[i]=='7'||dat$Month[i]=='8') {dat$Season[i]="Summer"}
  
if(dat$Month[i]=="9" || dat$Month[i]=='10'||dat$Month[i]=='11') {dat$Season[i]="Fall"}  
  
}

#melt the data frame for building the plots

dat.temp<-dat[,-c(1,2,22)]

dat.melt<-melt(dat.temp, id=c("Station_alt","Date","Year","Month","Day","Season"))
```
###Salinity
\
\
```{r,echo = F, warning = F, comment = F, message=F}
##########################

#subset the water column characteristics
dat.wc<-subset(dat.melt,variable==c("Temp_C","Salinity_ppt","Ni_Nitrate","P","Silica"))

sal<-dat.melt[dat.melt$variable=="Salinity_ppt",]
temp<-dat.melt[dat.melt$variable=="Temp_C",]
Nitrate<-dat.melt[dat.melt$variable=="Ni_Nitrate",]
Phos<-dat.melt[dat.melt$variable=="P",]
Silica<-dat.melt[dat.melt$variable=="Silica",]
Ammon<-dat.melt[dat.melt$variable=="NH4",]


###################
#Salinity by season
###################

sal_by_season<- sal %>% group_by(Station_alt,Year,Season)%>%summarize(SampleN=length(value),mean_var=mean(value,na.rm=T),sd_var=sd(value,na.rm = T))

sal_by_Year<- sal %>% group_by(Year,Season)%>%summarize(SampleN=length(value),mean_var=mean(value,na.rm=T),sd_var=sd(value,na.rm = T),min_var=min(value,na.rm=T), max_var=max(value,na.rm=T))

sal.plot<-ggplot(sal_by_season, aes(x=Station_alt,y=mean_var,group=Season))+
  geom_point(aes(col=Season,shape=Season),size = 3)+
  #scale_shape_manual(values=c(16,17,15,8))+
  geom_line(aes(col=Season),linetype = 1,size=1)+
  scale_color_grey(name="Season")+
  ylab("ppt")+
  xlab("Station ID")+
  cust_theme+
  ggtitle("Salinity")+
  facet_grid(Year~.)


#make a table
sal.table<-sal_by_Year %>%
  kable() %>%
   kable_styling(bootstrap_options = "striped", full_width = F)

sal.plot
sal.table

```
###Temperature
\ 
\
```{r,echo = F, warning = F, comment = F, message=F}

temp_by_season<- temp %>% group_by(Station_alt,Year,Season)%>%summarize(SampleN=length(value),mean_var=mean(value,na.rm=T),sd_var=sd(value,na.rm = T))

temp_by_Year<- temp %>% group_by(Year,Season)%>%summarize(SampleN=length(value),mean_var=mean(value,na.rm=T),sd_var=sd(value,na.rm = T),min_var=min(value,na.rm=T), max_var=max(value,na.rm=T))

temp.plot<-ggplot(temp_by_season, aes(x=Station_alt,y=mean_var,group=Season))+
  geom_point(aes(col=Season,shape=Season),size = 3)+
  #scale_shape_manual(values=c(16,17,15,8))+
  geom_line(aes(col=Season),linetype = 1,size=1)+
  scale_color_grey(name="Season")+
  ylab("deg C")+
  xlab("Station ID")+
  cust_theme+
  ggtitle("Temperature")+
  facet_grid(Year~.)

temp.plot
temp.table<-temp_by_Year %>%
  kable() %>%
   kable_styling(bootstrap_options = "striped", full_width = F)
temp.table
```
###Nitrate
\
\
```{r,echo = F, warning = F, comment = F, message=F}
###################
#Nitrate by season
###################


Nit_by_season<- Nitrate %>% group_by(Station_alt,Year,Season,variable)%>%summarize(SampleN=length(value),mean_var=mean(value,na.rm=T),sd_var=sd(value,na.rm = T))

#Tables to save
Nit_by_Year<- Nitrate %>% group_by(Year,Season)%>%summarize(SampleN=length(value),mean_var=mean(value,na.rm=T),sd_var=sd(value,na.rm = T),min_var=min(value,na.rm=T), max_var=max(value,na.rm=T))

Nit.plot<-ggplot(Nit_by_season, aes(x=Station_alt,y=mean_var,group=Season))+
  geom_point(aes(col=Season,shape=Season),size = 3)+
  #scale_shape_manual(values=c(16,17,15,8))+
  geom_line(aes(col=Season),linetype = 1,size=1)+
  scale_color_grey(name="Season")+
  ylab(expression(paste("Nitrate (",mu,"mol ",l^-1,")")))+
  xlab("Station ID")+
  cust_theme+
  ggtitle("Nitrate")+
  facet_grid(Year~.)

Nit.plot
Nit.table<-Nit_by_Year %>%
  kable() %>%
   kable_styling(bootstrap_options = "striped", full_width = F)
Nit.table

```
###Phosphate
\
\
```{r,echo = F, warning = F, comment = F, message=F}

###################
#P by season
###################
Phos_by_season<- Phos %>% group_by(Station_alt,Year,Season,variable)%>%summarize(SampleN=length(value),mean_var=mean(value,na.rm=T),sd_var=sd(value,na.rm = T))

#Tables to save
Phos_by_Year<- Phos %>% group_by(Year,Season)%>%summarize(SampleN=length(value),mean_var=mean(value,na.rm=T),sd_var=sd(value,na.rm = T),min_var=min(value,na.rm=T), max_var=max(value,na.rm=T))

Phos.plot<-ggplot(Phos_by_season, aes(x=Station_alt,y=mean_var,group=Season))+
  geom_point(aes(col=Season,shape=Season),size = 3)+
  #scale_shape_manual(values=c(16,17,15,8))+
  geom_line(aes(col=Season),linetype = 1,size=1)+
  scale_color_grey(name="Season")+
  ylab(expression(paste("Phosphate (",mu,"mol ",l^-1,")")))+
  xlab("Station ID")+
  cust_theme+
  ggtitle("Phosphate")+
  facet_grid(Year~.)

Phos.plot
Phos.table<-Phos_by_Year %>%
  kable() %>%
   kable_styling(bootstrap_options = "striped", full_width = F)
Phos.table

```
###Ammonium
\
\
```{r,echo = F, warning = F, comment = F, message=F}
###################
#Ammonium by season
###################
Ammon_by_season<- Ammon %>% group_by(Station_alt,Year,Season)%>%summarize(SampleN=length(value),mean_var=mean(value,na.rm=T),sd_var=sd(value,na.rm = T))

#Tables to save
Ammon_by_Year<- Ammon %>% group_by(Year,Season)%>%summarize(SampleN=length(value),mean_var=mean(value,na.rm=T),sd_var=sd(value,na.rm = T),min_var=min(value,na.rm=T), max_var=max(value,na.rm=T))

Ammon.plot<-ggplot(Ammon_by_season, aes(x=Station_alt,y=mean_var,group=Season))+
  geom_point(aes(col=Season,shape=Season),size = 3)+
  #scale_shape_manual(values=c(16,17,15,8))+
  geom_line(aes(col=Season),linetype = 1,size=1)+
  scale_color_grey(name="Season")+
  ylab(expression(paste("Ammonium (",mu,"mol ",l^-1,")")))+
  xlab("Station ID")+
  cust_theme+
  ggtitle("Phosphate")+
  facet_grid(Year~.)

Ammon.plot
Ammon.table<-Ammon_by_Year %>%
  kable() %>%
   kable_styling(bootstrap_options = "striped", full_width = F)

Ammon.table
```
###Silica
\
\
```{r,echo = F, warning = F, comment = F, message=F}
###################
#Silica by season
###################


Silica_by_season<- Silica %>% group_by(Station_alt,Year,Season)%>%summarize(SampleN=length(value),mean_var=mean(value,na.rm=T),sd_var=sd(value,na.rm = T))

Silica_by_Year<- Silica %>% group_by(Year,Season)%>%summarize(SampleN=length(value),mean_var=mean(value,na.rm=T),sd_var=sd(value,na.rm = T),min_var=min(value,na.rm=T), max_var=max(value,na.rm=T))

Silica.plot<-ggplot(Silica_by_season, aes(x=Station_alt,y=mean_var,group=Season))+
  geom_point(aes(col=Season,shape=Season),size = 3)+
  #scale_shape_manual(values=c(16,17,15,8))+
  geom_line(aes(col=Season),linetype = 1,size=1)+
  scale_color_grey(name="Season")+
  ylab(expression(paste("Silica (",mu,"mol ",l^-1,")")))+
  xlab("Station ID")+
  cust_theme+
  ggtitle("Silica")+
  facet_grid(Year~.)

Silica.plot

sil.table<-Silica_by_Year %>%
  kable() %>%
   kable_styling(bootstrap_options = "striped", full_width = F)
sil.table

```
\
Doing some statistics to match the plots...


